A Framework for Equivalence of Cast Calculi and a new Cast Representation

### 2020-10-14 paper.tex
mention de Bruijn indexes
shortened and revised the abstract, other minor edits to get page count down to 25

\begin{document}
%
\title{A Framework for Equivalence of Cast Calculi and a new Cast Representation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Kuang-Chen Lu\inst{1} \and
Jeremy G. Siek\inst{2} \and
Andre Kuhlenschmidt\inst{2}
}
%
\authorrunning{Lu et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Brown University, Providence RI 02912, USA \and
Indiana University, Bloomington IN 47405, USA}
% \institute{Princeton University, Princeton NJ 08544, USA \and
% Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
% \email{lncs@springer.com}\\
% \url{http://www.springer.com/gp/computer-science/lncs} \and
% ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
% \email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
% Background
Gradual typing combines dynamic typing and static typing in one
language to provide the advantages of both typing
disciplines. Implementing gradual typing with reasonable overhead,
however, is not trivial. Casts accumulate on higher-order values and
continuations, changing the overall space complexity. Several
space-efficient cast representations have been proposed, notably
coercions in normal forms and threesomes.
A correct cast representations shall be proved not changing the overall
behavior of language implementations except saving space.
Existing proofs typically go through a bisimulation between
the language semantics and a space-efficient one 
\citep{Siek:2010:TWB:1706299.1706342,Siek:2015:BCT:2737924.2737968}.
% Problem
The similarity of these proofs suggests that people should be able
to prove correctness for a cast representation by proving it holds
certain properties.
% Solution
We present an abstract data type, namely Cast ADT,
that captures interface functions a cast representations typically 
provides, and an abstract machine parameterized over Cast ADTs.
Our abstract machine comes with generic theorems,
% Why is our solution a step forward
with which the correctness proofs can be done by just proving 
properties of a specific Cast ADT.
% %
% To isolate cast representations from the remaining part of 
% implementations, we present a abstract data type, namely Cast ADT,
% that captures interface functions a cast representations typically 
% includes, and an abstract machine parameterized over Cast ADTs.
% %
% With our generic abstract machine, the correctness proofs can be
% done by proving properties of a specific Cast ADT, which is simpler
% than proofs regarding the entire dynamic semantics.
%%%%%
% %
% But they are presented in
% different dynamic semantics, which make it difficult to compare the
% key differences between the cast representations.
% %
% We present a
% abstract data type, namely Cast ADT, that captures interface functions
% a cast representations typically includes, and an abstract machine
% parameterized over Cast ADTs.
% %
% The Cast ADT brings another benefit:
% space-efficient cast representations shall be proved not changing the
% overall behavior of language implementations except saving space.
% %
% With our generic abstract machine, these proofs can be done by proving
% properties of a specific Cast ADT, which is simpler than proofs
% regarding the entire dynamic semantics.
\todo[inline]{JS: The stuff above about ``make it difficult to compare''
  is a weak argument. I would recommend deleting that argument and
  instead focus on the benefits regarding the proofs of correctness.}
\todo[inline]{KC: updated.}

We have successfully applied our new approach to the proofs of
correctness for two space-efficient cast representations from the
literature
\citep{Siek:2010:TWB:1706299.1706342,siek2012interpretations,Siek:2015:BCT:2737924.2737968}
and believe this framework is also applicable to other cast
representations
\citep{Garcia:2013:CTB:2500365.2500603,siek2012interpretations}.  We
also present a new cast representation, hypercoercions, that is more
friendly to mechanized meta-theory and consumes less memory than
coercions in normal forms. We prove correctness for two variants of
hypercoercions.

@@

\todo[inline]{KC: is this the right place to mention that in Agda
code we use de Bruijn indexes?}

@@

@@

\item If the middle is of base types, then the remaining 62 bits are
used to represent the head (51 bits), middle (10 bit), and tail (1
bit).
\begin{itemize}
\item The head requires 1 bit to differentiate between $\hche$
and $\hchp{P}{l}$ and then 50 bits could be used for the label $l$.
\item The middle would use 10 bits to differentiate all the base types.
\item The tail requires 1 bits to differentiate between $\hcte$ and $\hcti{P}$, 
where $P$ must be equal to the middle type.
\end{itemize}
The bits for blame labels represent an index into a table of blame information.
In the event that a program requires too many blame labels, then the 
implementation
can fallback to using the heap-allocated structure for more
coercions.
\end{itemize}

@@

We present the first framework for mechanized proofs of correctness
for space-efficient cast representations.
%
The framework includes abstract data types for cast representations
(Cast ADT and its refinements Lazy D Cast ADT and \LUD\ \CastADT), a
parameterized abstract machine \SMachine{C}, a theorem stating that
\SMachine{C} is equivalent to \LDMachine\ when $C$ is a Lazy D Cast
ADT, and a theorem which states that \SMachine{C} is equivalent to
\LUDMachine\ when $C$ is a Lazy UD Cast ADT.
%
%The Lazy D Cast ADT refines the Cast ADT. 
%
We applied this framework to Lazy UD coercions in normal form, Lazy UD
threesomes, and hypercoercions and expect that this framework is also
applicable to many other cast representations.

>>>>

\begin{document}
%
\title{A Framework for Equivalence of Cast Calculi and a new Cast Representation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Kuang-Chen Lu\inst{1} \and
Jeremy G. Siek\inst{2} \and
Andre Kuhlenschmidt\inst{2}
}
%
\authorrunning{Lu et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Brown University, Providence RI 02912, USA \and
Indiana University, Bloomington IN 47405, USA}
% \institute{Princeton University, Princeton NJ 08544, USA \and
% Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
% \email{lncs@springer.com}\\
% \url{http://www.springer.com/gp/computer-science/lncs} \and
% ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
% \email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
% Background
Gradual typing combines dynamic typing and static typing in one
language to provide the advantages of both typing
disciplines. Implementing gradual typing with reasonable overhead,
however, is not trivial. Casts accumulate on higher-order values and
continuations, changing the overall space complexity. Several
space-efficient cast representations have been proposed, notably
coercions in normal forms and threesomes.
% Problem
The proofs of correctness of these cast representations are
challenging, involving a bisimulation between the language semantics
and a space-efficient implementation
\citep{Siek:2010:TWB:1706299.1706342,Siek:2015:BCT:2737924.2737968}.
The prospect of mechanizing these proofs is even more daunting.
% Solution
In this paper we present a framework in Agda that reduces the cost of
mechanized proofs of correctness for cast representations.

%% The abstract is too long... I wrote the following but I'm now
%% cutting it. No need for that much detail here. -Jeremy
%% We present
%% an abstract data type, the Cast ADT, that defines an interface for
%% cast representations and we specify which properties are necessary for
%% a Cast ADT to be correct. We then define an abstract machine
%% parameterized over Cast ADTs and prove its correctness.  A user of our
%% framework proves the correctness of their cast representation by
%% implementing the Cast ADT, proving that it satisfies the correctness
%% properties, and then instantiating our generic theorem.

%% A correct cast representations shall be proved not changing the overall
%% behavior of language implementations except saving space.
%% Existing proofs typically go through a 
%% % Problem
%% The similarity of these proofs suggests that people should be able
%% to prove correctness for a cast representation by proving it holds
%% certain properties.
%% % Solution
%% We present an abstract data type, namely Cast ADT,
%% that captures interface functions a cast representations typically 
%% provides, and an abstract machine parameterized over Cast ADTs.
%% Our abstract machine comes with generic theorems,
%% % Why is our solution a step forward
%% with which the correctness proofs can be done by just proving 
%% properties of a specific Cast ADT.
% %
% To isolate cast representations from the remaining part of 
% implementations, we present a abstract data type, namely Cast ADT,
% that captures interface functions a cast representations typically 
% includes, and an abstract machine parameterized over Cast ADTs.
% %
% With our generic abstract machine, the correctness proofs can be
% done by proving properties of a specific Cast ADT, which is simpler
% than proofs regarding the entire dynamic semantics.
%%%%%
% %
% But they are presented in
% different dynamic semantics, which make it difficult to compare the
% key differences between the cast representations.
% %
% We present a
% abstract data type, namely Cast ADT, that captures interface functions
% a cast representations typically includes, and an abstract machine
% parameterized over Cast ADTs.
% %
% The Cast ADT brings another benefit:
% space-efficient cast representations shall be proved not changing the
% overall behavior of language implementations except saving space.
% %
% With our generic abstract machine, these proofs can be done by proving
% properties of a specific Cast ADT, which is simpler than proofs
% regarding the entire dynamic semantics.
\todo[inline]{JS: The stuff above about ``make it difficult to compare''
  is a weak argument. I would recommend deleting that argument and
  instead focus on the benefits regarding the proofs of correctness.}
\todo[inline]{KC: updated.}

We apply our framework to obtain proofs of correctness for two
space-efficient cast representations from the literature
\citep{Siek:2010:TWB:1706299.1706342,siek2012interpretations,Siek:2015:BCT:2737924.2737968}
and believe this framework is also applicable to other cast
representations
\citep{Garcia:2013:CTB:2500365.2500603,siek2012interpretations}.
Finally, we present a new cast representation, named hypercoercions,
that is more friendly to mechanization and consumes less memory than
coercions in normal forms. We prove correctness for two variants of
hypercoercions using our framework.

@@

@@

\todo[inline]{KC: is here the right place to mention that in Agda
code we use de Bruijn indexes?}
In the Agda code we use de Bruijn indices instead of named variables.

@@

\item If the middle is of base types, then the remaining 62 bits are
used to represent the head (51 bits), middle (10 bit), and tail (1
bit).
\begin{itemize}
\item The head requires 1 bit to differentiate between $\hche$
and $\hchp{P}{l}$ and then 50 bits could be used for the label $l$.
\item The middle would use 10 bits to differentiate all the base types.
\item The tail requires 1 bits to differentiate between $\hcte$ and $\hcti{P}$, 
where $P$ must be equal to the middle type.
\end{itemize}
The bits for blame labels represent an index into a table of blame
information.  If a program requires too many blame labels, then the
implementation can fallback to using the heap-allocated structure for
more coercions.
\end{itemize}

@@

We present the first framework for mechanized proofs of correctness
for space-efficient cast representations.
%
The framework includes abstract data types for cast representations,
%(Cast ADT and its refinements Lazy D Cast ADT and \LUD\ \CastADT)
a parameterized abstract machine \SMachine{C}, a theorem stating that
\SMachine{C} is equivalent to \LDMachine\ when $C$ is a Lazy D Cast
ADT, and a theorem which states that \SMachine{C} is equivalent to
\LUDMachine\ when $C$ is a Lazy UD Cast ADT.
%
%The Lazy D Cast ADT refines the Cast ADT. 
%
We applied this framework to Lazy UD coercions in normal form, Lazy UD
threesomes, and hypercoercions and expect that this framework is also
applicable to many other cast representations.
